Project Part 3

### 2024-04-05 code/data_analysis_code.qmd
Updated the README file for running code
Project Part 4 complete
Done

List of packages needed for the whole data processing/data analysis

```{r}
library(here)
library(readxl)
library(tidyverse)
library(naniar)
library(vtable)
library(corrplot)
library(RColorBrewer)
library(ggpubr)
library(rstatix)
```

##Data Cleaning

### HAPIN Guatemala data cleaning

```{r}
#Load the files and assign them to independent data frames
hapin <- read_csv(here("data", "raw-data", "HAPIN_EX17_Guatemala_20230728.csv"))
xrflist <- read_csv(here("data", "raw-data", "sa_filters_emollinedo.csv"))

#Select only the variables I am interested to work with
hapin <- hapin %>% select(trt_blinded, timepoint, h41_m_ecm_fid, h41_b_ecm_fid, h41b_filter1, h41b_filter2, h42_cook, h42_smoke, h42_coil, h42_trash, h42_kero, h42_incence, h42_generator, h42_smoked, h42_crop, h42_smoky_kitc, ECM_grav_neph_conc_M, ECM_bc_conc_M, CO_avg_ppm_M)

#Make sure all filter IDs are upper case, using the `mutate()` function with `toupper()`
hapin <- hapin %>% 
  mutate(h41_m_ecm_fid = toupper(h41_m_ecm_fid),
         h41b_filter1 = toupper(h41b_filter1),
         h41_b_ecm_fid = toupper(h41_b_ecm_fid),
         h41b_filter2 = toupper(h41b_filter2))

#Select all filters that match with the list of filters from XRF, using the `filter()` function.
hapin <- hapin %>%
  filter(h41_m_ecm_fid %in% xrflist$filter_id |
           h41b_filter1 %in% xrflist$filter_id |
           h41_b_ecm_fid %in% xrflist$filter_id |
           h41b_filter2 %in% xrflist$filter_id)

#Concatenate `trt_blinded` and `timepoint` into a new temporary variable
hapin$concatenated <- paste(hapin$trt_blinded, hapin$timepoint)

#Create a new variable `fueltype` that categorizes all the observations by type of fuel ('R' means the visit belongs to the 'Control' treatment and 'Q' for the 'Intervention' treatment)
hapin <- mutate(hapin, fueltype = case_when(
  endsWith(concatenated, "BL") ~ "Biomass", #BL corresponds to the baseline, where all the participants had a biomass type fuel stove
  endsWith(concatenated, "R P1") ~ "Biomass",
  endsWith(concatenated, "R P2") ~ "Biomass",
  endsWith(concatenated, "Q P1") ~ "LPG",
  endsWith(concatenated, "Q P2") ~ "LPG",
  endsWith(concatenated, "R BLP1") ~ "Biomass",
  endsWith(concatenated, "R P1P2") ~ "Biomass",
  endsWith(concatenated, "R P2B1") ~ "Biomass",
  endsWith(concatenated, "Q BLP1") ~ "LPG",
  endsWith(concatenated, "Q P1P2") ~ "LPG",
  endsWith(concatenated, "Q P2B1") ~ "LPG"))

#Rename some of the variables and delete unnecessary ones
hapin <- hapin %>% rename(filter_id = "h41_m_ecm_fid", dup1_filterid = "h41b_filter1", dup2_filterid = "h41b_filter2",
                         blank_filterid = "h41_b_ecm_fid", arm = "trt_blinded", pm25 = "ECM_grav_neph_conc_M",
                         bc = "ECM_bc_conc_M", co = "CO_avg_ppm_M", stove = "h42_cook", smoke = "h42_smoke",
                         coil = "h42_coil", trash = "h42_trash", kerosene = "h42_kero", incense = "h42_incence",                                   generator = "h42_generator", smoky = "h42_smoked", crop = "h42_crop", stove_other = "h42_smoky_kitc") %>% 
  select(-c(concatenated, timepoint)) %>% 
  mutate(arm = case_when(arm == 'R' ~ 'Control', arm == 'Q' ~ 'Intervention', TRUE ~ arm)) #Recode values from the 'arm' variable

#Creating the dataframe for personal sample filters, and deleting the other columns with unnecessary filter IDs
hapin_personal <- hapin %>% filter(filter_id %in% xrflist$filter_id) %>% 
  select(-c(blank_filterid, dup1_filterid, dup2_filterid))

#Creating the dataframe for duplicates in column 4, and deleting the other columns with unnecessary filter IDs, and renaming `dup1_filterid` to `filter_id`.
hapin_dup1 <- hapin %>% filter(dup1_filterid %in% xrflist$filter_id) %>% 
  select(-c(blank_filterid, filter_id, dup2_filterid)) %>% 
  rename(filter_id = "dup1_filterid")

#Creating the dataframe for duplicates in column 5, and deleting the other columns with unnecessary filter IDs, and renaming `dup2_filterid` to `filter_id`.
hapin_dup2 <- hapin %>% filter(dup2_filterid %in% xrflist$filter_id) %>% 
  select(-c(blank_filterid, filter_id, dup1_filterid)) %>% 
  rename(filter_id = "dup2_filterid")

#Creating the dataframe for blank filters, and deleting the other columns with unnecessary filter IDs, and renaming `blank_filterid` to `filter_id`.
hapin_blanks <- hapin %>% filter(blank_filterid %in% xrflist$filter_id) %>% 
  select(-c(filter_id, dup1_filterid, dup2_filterid)) %>% 
  rename(filter_id = "blank_filterid")

#Combining the `personal` and `dup` dataframes into a single one, using `bind_rows()`
hapin_filters <- bind_rows(hapin_personal, hapin_dup1, hapin_dup2)

#Save the samples and blanks dataframes to .RDS format files
saveRDS(hapin_filters, file = here("data", "processed-data", "hapin_samples.rds"))
saveRDS(hapin_blanks, file = here("data", "processed-data", "hapin_blanks.rds"))
```

### Chemical species Concentrations data cleaning

```{r}
#Loading the chemical species concentration data
concentration <- read_excel(here("data", "raw-data", "conc_unc_hapingt_emollinedo.xlsx"), sheet = "Concentration")

#Loading the chemical species limit of detection values
lod <- read_excel(here("data", "raw-data", "conc_unc_hapingt_emollinedo.xlsx"), sheet = "DL")

#having a glimpse of this dataframe
glimpse(concentration)

#Remove brackets, convert to numeric, and round to 4 decimal places
concentration <- concentration %>%
  mutate(across(3:26, ~ round(as.numeric(str_remove_all(., "\\[|\\]")), 4))) %>% 
  rename(filter_id = "Filter")

#Create dataframe with the concentrations for the blank filters for each element
blanks <- concentration %>% filter(filter_id %in% hapin_blanks$filter_id) %>% 
  select(-c(filter_id, Type))

#Remove blank filters from the `concentration` dataframe
concentration <- concentration %>%
  anti_join(hapin_blanks, by = "filter_id")

#Calculate the median for each element, using the blank filter values
median_blanks <- blanks %>% select(Mg: Na) %>%
  map_dbl(median) %>% round(4)

median_blanks

#Round the limit of detection values to 4 decimal places
lod <- lod %>% mutate(`Concentration (ug/cm2)` = round(`Concentration (ug/cm2)`, 4))
print(lod)

#Estimate the percentage of values below detection limit for each element
bdl <- concentration %>%
  summarise(
    percent_bdl_mg = sum(Mg < 0.0051) / n() * 100,
    percent_bdl_al = sum(Al < 0.0059) / n() * 100,
    percent_bdl_si = sum(Si < 0.0068) / n() * 100,
    percent_bdl_s = sum(S < 0.0026) / n() * 100,
    percent_bdl_k = sum(K < 0.0040) / n() * 100,
    percent_bdl_ca = sum(Ca < 0.0040) / n() * 100,
    percent_bdl_ti = sum(Ti < 0.0086) / n() * 100,
    percent_bdl_cr = sum(Cr < 0.0055) / n() * 100,
    percent_bdl_mn = sum(Mn < 0.0030) / n() * 100,
    percent_bdl_fe = sum(Fe < 0.0043) / n() * 100,
    percent_bdl_ni = sum(Ni < 0.0043) / n() * 100,
    percent_bdl_cu = sum(Cu < 0.0080) / n() * 100,
    percent_bdl_zn = sum(Zn < 0.0024) / n() * 100,
    percent_bdl_ga = sum(Ga < 0.0049) / n() * 100,
    percent_bdl_as = sum(As < 0.0062) / n() * 100,
    percent_bdl_se = sum(Se < 0.0181) / n() * 100,
    percent_bdl_cd = sum(Cd < 0.0932) / n() * 100,
    percent_bdl_in = sum(In < 0.1089) / n() * 100,
    percent_bdl_sn = sum(Sn < 0.1844) / n() * 100,
    percent_bdl_te = sum(Te < 0.2166) / n() * 100,
    percent_bdl_i = sum(I < 0.3149) / n() * 100,
    percent_bdl_pb = sum(Pb < 0.0324) / n() * 100,)

print(bdl)

#Select only the filter ID, type and the 10 elements that have more than 50% of their data not below the detection limit
concentration <- concentration %>% select(c(filter_id, Type, Mg, Al, Si, S, K, Ca, Ti, Mn, Fe, Zn)) %>% 
  arrange(filter_id) #Also I decided to arrange them ascending based on the filter_id

#Substract the median blank value for the elements that need blank adjustment
concentration <- concentration %>%
  mutate(Mg = Mg - 0.0112, S = S - 0.0044, K = K - 0.0014, Ca = Ca - 0.0048, Fe = Fe - 0.0038, Zn = Zn - 0.0045)

#Replace all values below the detection limit with DL/sqrt(2)
concentration <- concentration %>% 
    mutate(Mg = round(if_else(Mg < 0.0051, 0.0051 / sqrt(2), Mg), 4), Al = round(if_else(Al < 0.0059, 0.0059 / sqrt(2), Al), 4),
           Si = round(if_else(Si < 0.0068, 0.0068 / sqrt(2), Si), 4), S = round(if_else(S < 0.0026, 0.0026 / sqrt(2), S), 4),
           K = round(if_else(K < 0.0016, 0.0016 / sqrt(2), K), 4), Ca = round(if_else(Ca < 0.0040, 0.0040 / sqrt(2), Ca), 4),
           Ti = round(if_else(Ti < 0.0086, 0.0086 / sqrt(2), Ti), 4), Mn = round(if_else(Mn < 0.0030, 0.0030 / sqrt(2), Mn), 4),
           Fe = round(if_else(Fe < 0.0043, 0.0043 / sqrt(2), Fe), 4), Zn = round(if_else(Zn < 0.0024, 0.0024 / sqrt(2), Zn), 4))

#I will save these concentrations in a new dataframe `concentrations_ugcm2` that will be used for uncertainty estimations in future steps
concentrations_ugcm2 <- concentration %>% select(-Type)

#Converting the concentrations from (ug/cm2) to (ug/m3)
concentration <- concentration %>%
  mutate(across(Mg:Zn, ~ round(.x * 0.7853 / 0.43, 4)))

#Now adding the BC concentrations value from the `hapin_filters` df
hapin_filters <- hapin_filters %>% arrange(filter_id) #First arrange in ascending order by filter_id
concentration <- bind_cols(concentration, hapin_filters$bc) %>% rename(BC = "...13") %>% select(-Type) #Join the BC concentrations to the `concentrations` df, also renaming the column to "BC" and deleting the 'Type' column

#Add the `arm` and `fueltype` variables to the `concentration` dataframe
concentration <- concentration %>%
  mutate(arm = hapin_filters$arm, fueltype = hapin_filters$fueltype)

#Check if there are any missing values from the concentration file
gg_miss_var(concentration)

#Use na.omit() to remove the spotted observation with missing value
concentration <- na.omit(concentration)

#Save this dataframe as RDS file
saveRDS(concentration, file = here("data", "processed-data", "concentration.rds"))
```

### Chemical species Uncertainties data processing

```{r}
#Load the uncertainty data
uncertainty <- read_excel(here("data", "raw-data", "conc_unc_hapingt_emollinedo.xlsx"), sheet = "Uncertainty")

#Remove the chemical species with more than 50% of the observations with values BDL
uncertainty <- uncertainty %>% select(c(Filter, Mg, Al, Si, S, K, Ca, Ti, Mn, Fe, Zn)) %>% #Select only the necessary variables
  rename(filter_id = "Filter") %>% #Rename the Filter column
  mutate_at(vars(Mg:Zn), ~ round(., 4)) %>% #Round the uncertainties to 4 decimal places
  arrange(filter_id) %>% #Arrange observations based on Filter ID
  anti_join(hapin_blanks, by = "filter_id") #Remove blank filters

#Have a glimpse of the data
glimpse(uncertainty)

#Create a list for the detection limits for each element
dl <- c(Mg = 0.0051, Al = 0.0059, Si = 0.0068, S = 0.0026, K = 0.0016, 
        Ca = 0.0040, Ti = 0.0086, Mn = 0.0030, Fe = 0.0043, Zn = 0.0024)

# Recalculate the values in df based on the conditions
uncertainty <- uncertainty %>%
  mutate(across(all_of(names(dl)),
                ~ round(if_else(concentrations_ugcm2[[cur_column()]] < dl[[cur_column()]], (5/6) * dl[[cur_column()]], .), 4),
                .names = "{.col}"))

# Define the variables to be recalculated
variables <- c('Mg', 'Al', 'Si', 'S', 'K', 'Ca', 'Ti', 'Mn', 'Fe', 'Zn')

# Recalculate the values using the given equation
uncertainty <- uncertainty %>%
  mutate(across(all_of(variables), ~ round( #Here, selecting all the variables to be re-calculated
    sqrt((0.7853/0.43)^2 * .^2 + (concentrations_ugcm2[[cur_column()]]/0.43)^2 * 0.05^2 + (-concentrations_ugcm2[[cur_column()]]*0.7853/0.43^2)^2 * (0.05 * 0.43)^2), 4), .names = "{.col}")) #Input the equation

#Load file to estimate BC uncertainties
atn <- read_csv(here("data", "raw-data", "ATN_summary.csv"))

#Explore the structure of the file
glimpse(atn)

#Estimate total attenuation by sample
atn <- atn %>% mutate(ATN = round(log(post_ATN/pre_ATN) *100, 2))

#I will remove the observation with `filter_id` 3M53864 from the `uncertainty` df, since this observation does not have a BC concentration available
uncertainty <- uncertainty %>% filter(filter_id != "3M53864") %>% 
  mutate(BC = round((1370 * sqrt((atn$ATN)^2 * 0.000005^2 + 0.000079^2 * 1.16^2)), 4))#Estimate the BC uncertainties

#I will remove the observation with `filter_id` 3M53864 from the `hapin_filters` df
hapin_filters <- hapin_filters %>% filter(filter_id != "3M53864")

#Add the `arm` and `fueltype` variables to the `uncertainty` dataframe
uncertainty <- uncertainty %>%
  mutate(arm = hapin_filters$arm, fueltype = hapin_filters$fueltype)

#Save this dataframe as RDS file
saveRDS(uncertainty, file = here("data", "processed-data", "uncertainty.rds"))
```

## Data Exploring

First, I produced two boxplots, that summarize the concentrations categorized by `arm` and `fueltype`. First I will create a dataframe in longer format using `pivot_longer()`. Then creating the plots using `ggplot()`. In addition, I will add the significance test value to the plots to visualize which chemical species are statistically different among groups.

```{r}
#Pivot longer the `concentration` df to make the boxplots
conc_longer <- concentration %>% #First a df with the `arm` variable
  pivot_longer(cols = -c(filter_id, fueltype, arm),
               names_to = "species", values_to = "concentration")

#conduct the t-test of concentration-fueltype
stat.test <- conc_longer %>% 
  group_by(species) %>% 
  t_test(concentration ~ fueltype) %>% 
  add_significance() #to add the significance values

#Add the position coordinates
stat.test <- stat.test %>% 
  add_x_position(x= "species", dodge = 0.8) %>% #x position is species, and the length of significance bar is 0.8
  rstatix::add_y_position(y.trans = function(x){log10(x)}) #add y position, in this case specify using a function that it is in log10 scale

#Creating the boxplot of concentrations by fueltype
options(scipen = 999) #to remove scientific notation for smaller values (i.e. changes from 1*10-3 to 0.001)
ggplot(conc_longer)+
  geom_boxplot(aes(x= species, y= concentration, fill= fueltype))+
  scale_fill_brewer(palette = "Paired")+
  labs(x= "Chemical species", y= "Concentration (ug/m3)")+
  scale_y_log10()+
  guides(fill= guide_legend(title= "Fuel type"))+ #Change legend from fuel_type to Fuel type
  theme_bw(base_size = 15)+
  stat_pvalue_manual(stat.test, label = "p.signif", tip.length = 0.001) #Add the significance bars, use 'p.signif' to add *** or 'ns'

#Creating the boxplot of concentrations by arm
options(scipen = 999) #to remove scientific notation for smaller values (i.e. changes from 1*10-3 to 0.001)
ggplot(conc_longer)+
  geom_boxplot(aes(x= species, y= concentration, fill= arm))+
  scale_fill_brewer(palette = "Paired")+
  labs(x= "Chemical species", y= "Concentration (ug/m3)")+
  scale_y_log10()+
  guides(fill= guide_legend(title= "Study Arm"))+ #Change legend from fuel_type to Fuel type
  theme_bw(base_size = 15)+
  stat_pvalue_manual(stat.test, label = "p.signif", tip.length = 0.001) #Add the significance bars, use 'p.signif' to add *** or 'ns'
```

The boxplots for `fueltype` and `arm` show what I was expecting in terms of how the data is distributed. The concentrations seem higher in some elemental especies such as BC, Ca, K, Mg, Mn and S. The following steps will be to evaluate if those differences in concentrations are statistically different.

And now producing summary statistics by type of fuel, using the `sumtable()` function.

```{r}
#Produce summary statistics by type of fuel
#Remove arm and fueltype columns
conc <- concentration %>% select(-arm)

#Using the vtable package, which creates a table
sumtable(conc, group = 'fueltype', group.long = T, add.median = T)
```

Now I created a correlation plot using the `corrplot()` function, to explore which elements could be more associated between each other.

@@

#Perform the spearman correlation test
cor <- stats::cor(cor, method = "pearson", use = "complete.obs") #I had to input the `use=` argument since there are missing values in the BC variable. I will correct them later.

#Create the correlation plot
corrplot(cor, method = "color", type = "lower", #Color form and displays at the lower portion
         col = COL2("RdYlBu", 20), order = "hclust", #Color palette, and order or the species
         tl.col = "orangered4", tl.srt = 0, tl.cex = 0.9,  #Color of the axis, position and size
         addCoef.col = "gray10", number.font = 2, number.cex = 0.8, #Display correlation number, set the font and size
         col.lim = c(0, 1), is.corr = T) #Set the color limits at the bar
```

The correlation plot above shows that Si and Fe are strongly correlated, which might suggest that these elements represent one specific source of pollution. Other high correlations observed are between Al and Si, and Fe and Ti.

I produced a summary table of the uncertainties, using the `vtable` package.

```{r}
#Remove the `arm` variable
unc_fuel <- uncertainty %>% select(-arm)

#Use the `sumtable()` function from the `vtable` package, which creates a table
sumtable(unc_fuel, group = 'fueltype', group.long = T, add.median = T, out = 'kable')

#Remove the `fueltype` variable
unc_arm <- uncertainty %>% select(-fueltype)

#Use the `sumtable()` function from the `vtable` package, which creates a table
sumtable(unc_arm, group = 'arm', group.long = T, add.median = T, out = 'kable')

```

The tables above are just a way to represent the summary statistics of the uncertainties, however, the uncertainties will only be useful for the PMF analysis after the conclusion of this project. One thing noticed also from this table, is that there are some infinite values from the BC concentrations, which also show that there are missing values that have to be corrected in earlier steps. From this tables we can also see that there are 376 observations from the 'Biomass' and 254 from the 'LPG' fueltypes. Meanwhile, there are 323 observations from the 'Control' arm and 307 from the 'Intervention' arm.

Now, checking if there are missing values in the `hapin` dataframe for any of the observations using the `gg_miss_var()` function.

```{r}
gg_miss_var(hapin)
```

As seen above, there are some missing values from the observational exposures (kerosene to coil), there is just one missing observation for PM2.5 and BC concentrations and there is more than 30 missing observations for CO (carbon monoxide). The missing observation of BC might explain the missing value spotted before, so this observation might actually have to be dropped down.

Now I created a barplot showing the distribution of responses for the categorical variables `kerosene`, `trash`, `stove_other`, `stove`, `smoky`, `incense`, `generator`, `smoke`, `crop`, and `coil`. First doing data manipulation so the df is useful for plotting.

```{r}
#Pivot longer the `hapin` df to make a barplot
hapin_longer <- hapin %>%
  select(c(`kerosene`, `trash`, `stove_other`, `stove`, `smoky`, `incense`, `generator`, `smoke`, `crop`, `coil`)) %>% #Here selecting only the necessary variables
  pivot_longer(cols = everything(),
               names_to = "exposure", values_to = "response") #Create the longer df

#Create a simple barplot
ggplot(hapin_longer, aes(x= exposure, fill= response))+
  geom_bar(position = "dodge")+
  theme_classic()
```

The bar plot shows that the majority of participants recorded to not being exposed to most of the exposures, except to `stove`, which is the stove that the participants use. Others also recorded that were exposed to other stoves, or to trash burning.

And finally plotting the PM2.5, BC and CO concentrations in different graphs using `ggplot()`

```{r}
#Histogram plot for the PM2.5 concentrations
ggplot(hapin, aes(x= pm25, fill= fueltype))+
  geom_histogram()+
  theme_classic()+
  labs(x= "PM2.5 Concentration (ug/m3)")

#Histogram plot for the BC concentrations
ggplot(hapin, aes(x= bc, fill= fueltype))+
  geom_histogram()+
  theme_classic()+
  labs(x= "Black Carbon Concentration (ug/m3)")

#Histogram plot for the CO concentrations
ggplot(hapin, aes(x= co, fill= fueltype))+
  geom_histogram()+
  theme_classic()+
  labs(x= "Carbon Monoxide Concentration (ppm)")
```

The histograms for the concentration of the three pollutants show that they don't follow a normal distribution. Most of the concentrations are low, and it is also seen that overall the concentrations in the LPG fueltype are lower than the Biomass group. If these variables will be used for modeling purposes, they should be log-transformed or using a logistic regression approach.

```{r}
#Create the correlation plot
corrplot(cor, method = "color", type = "lower", #Color form and displays at the lower portion
         col = COL2("RdYlBu", 20), order = "hclust", #Color palette, and order or the species
         tl.col = "orangered4", tl.srt = 0, tl.cex = 1,  #Color of the axis, position and size
         addCoef.col = "gray10", number.font = 2, number.cex = 0.8, #Display correlation number, set the font and size
         col.lim = c(0, 1), is.corr = T) #Set the color limits at the bar
```

@@

To answer question 3, I will conduct a regression analysis. But first I will merge the data from `hapin_filters` with the `concentration` data.

@@

@@

```{r}
#First, loading the `tidymodels` package
library(tidymodels)

# Create a generalized linear regression model specification using gamma distribution (This will be used for all the models)
model_spec <- linear_reg(mode = "regression") %>%
  set_engine("glm", family = Gamma(link = "log"))

@@

# Create a workflow
workflowBC <- workflow() %>%
  add_recipe(recipeBC) %>%
  add_model(model_spec)

# Fit the model to the data
fitBC <- workflowBC %>%
  fit(data = hapin_merged)

@@

# Create a workflow
workflowCa <- workflow() %>%
  add_recipe(recipeCa) %>%
  add_model(model_spec)

# Fit the model to the data
fitCa <- workflowCa %>%
  fit(data = hapin_merged)

@@

@@

# Create a workflow
workflowK <- workflow() %>%
  add_recipe(recipeK) %>%
  add_model(model_spec)

# Fit the model to the data
fitK <- workflowK %>%
  fit(data = hapin_merged)

@@

# Create a workflow
workflowMg <- workflow() %>%
  add_recipe(recipeMg) %>%
  add_model(model_spec)

# Fit the model to the data
fitMg <- workflowMg %>%
  fit(data = hapin_merged)

@@

# Create a workflow
workflowMn <- workflow() %>%
  add_recipe(recipeMn) %>%
  add_model(model_spec)

# Fit the model to the data
fitMn <- workflowMn %>%
  fit(data = hapin_merged)

@@

# Create a workflow
workflowSul <- workflow() %>%
  add_recipe(recipeSul) %>%
  add_model(model_spec)

# Fit the model to the data
fitSul <- workflowSul %>%
  fit(data = hapin_merged)

@@

# Create a workflow
workflowAl <- workflow() %>%
  add_recipe(recipeAl) %>%
  add_model(model_spec)

# Fit the model to the data
fitAl <- workflowAl %>%
  fit(data = hapin_merged)

# Create a workflow
workflowFe <- workflow() %>%
  add_recipe(recipeFe) %>%
  add_model(model_spec)

# Fit the model to the data
fitFe <- workflowFe %>%
  fit(data = hapin_merged)

@@

# Create a workflow
workflowSi <- workflow() %>%
  add_recipe(recipeSi) %>%
  add_model(model_spec)

# Fit the model to the data
fitSi <- workflowSi %>%
  fit(data = hapin_merged)

@@

# Create a workflow
workflowTi <- workflow() %>%
  add_recipe(recipeTi) %>%
  add_model(model_spec)

# Fit the model to the data
fitTi <- workflowTi %>%
  fit(data = hapin_merged)

@@

# Create a workflow
workflowZn <- workflow() %>%
  add_recipe(recipeZn) %>%
  add_model(model_spec)

# Fit the model to the data
fitZn <- workflowZn %>%
  fit(data = hapin_merged)

@@

Based on the previous models, it seems like the most important variables are `fueltype` (as expected), `kerosene`, `stove_other` and `smoke` depending on the chemical species. So I will test new regression models for all species (except Al and Zn, since they seem not to be explained by any of the exposures), using only their most significant predictors.

```{r}
###########To predict the BC variable##############
# Create the recipe for preprocessing
recipeBC_red <- recipe(BC ~ kerosene + stove_other + fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowBC_red <- workflow() %>%
  add_recipe(recipeBC_red) %>%
  add_model(model_spec)

# Fit the model to the data
fitBC_red <- workflowBC_red %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsBC_red <- fitBC_red %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsBC_red)

###########To predict the Ca variable##############
# Create the recipe for preprocessing
recipeCa_red <- recipe(Ca ~ fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowCa_red <- workflow() %>%
  add_recipe(recipeCa_red) %>%
  add_model(model_spec)

# Fit the model to the data
fitCa_red <- workflowCa_red %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsCa_red <- fitCa_red %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsCa_red)

###########To predict the K variable##############
# Create the recipe for preprocessing
recipeK_red <- recipe(K ~ smoke + stove_other + fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowK_red <- workflow() %>%
  add_recipe(recipeK_red) %>%
  add_model(model_spec)

# Fit the model to the data
fitK_red <- workflowK_red %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsK_red <- fitK_red %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsK_red)

###########To predict the Mg variable##############
# Create the recipe for preprocessing
recipeMg_red <- recipe(Mg ~ kerosene + fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowMg_red <- workflow() %>%
  add_recipe(recipeMg_red) %>%
  add_model(model_spec)

# Fit the model to the data
fitMg_red <- workflowMg_red %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsMg_red <- fitMg_red %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsMg_red)

###########To predict the Mn variable##############
# Create the recipe for preprocessing
recipeMn_red <- recipe(Mn ~ fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowMn_red <- workflow() %>%
  add_recipe(recipeMn_red) %>%
  add_model(model_spec)

# Fit the model to the data
fitMn_red <- workflowMn_red %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsMn_red <- fitMn_red %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsMn_red)

###########To predict the S variable##############
# Create the recipe for preprocessing
recipeS_red <- recipe(S ~ fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowS_red <- workflow() %>%
  add_recipe(recipeS_red) %>%
  add_model(model_spec)

>>>>

List of packages needed for the whole data processing and data analysis

```{r}
library(here)
library(readxl)
library(tidyverse)
library(corrplot)
library(RColorBrewer)
library(ggpubr)
library(rstatix)
library(tidymodels)
```

Open the `concentration` and `hapin_samples` .RDS files to conduct the analyses.

```{r}
concentration <- read_rds(here("data", "processed-data", "concentration.rds"))
hapin_filters <- read_rds(here("data", "processed-data", "hapin_samples.rds"))
```

@@

#Perform the spearman correlation test
cor <- stats::cor(cor, method = "pearson")

#Create the correlation plot
corrplot(cor, method = "color", type = "lower", #Color form and displays at the lower portion
         col = COL2("RdYlBu", 20), order = "hclust", #Color palette, and order or the species
         tl.col = "orangered4", tl.srt = 0, tl.cex = 1,  #Color of the axis, position and size
         addCoef.col = "gray10", number.font = 2, number.cex = 0.8, #Display correlation number, set the font and size
         col.lim = c(0, 1), is.corr = T) #Set the color limits at the bar
```

@@

To answer question 3, I will conduct a regression analysis. But first I will merge the data from `hapin_filters` with the `concentration` data. Modeling can be difficult if there are missing values, so I will additionally clean the `hapin_merged` df, so it removes all NAs.

@@

#Remove pm25, bc and co, and additionally drop all NAs
hapin_merged <- hapin_merged %>%
  select(-c(pm25, bc, co)) %>%
  drop_na()

@@

```{r}
# Create a generalized linear regression model specification using gamma distribution (This will be used for all the models)
gamma_spec <- linear_reg(mode = "regression") %>%
  set_engine("glm", family = Gamma(link = "log"))

#Fit the model
fitAl <- workflow() %>%
  add_recipe(recipeAl) %>%
  add_model(gamma_spec) %>%
  fit(data = hapin_merged)

@@

#Fit the model
fitBC <- workflow() %>%
  add_recipe(recipeBC) %>%
  add_model(gamma_spec) %>%
  fit(data = hapin_merged)

@@

#Fit the model
fitCa <- workflow() %>%
  add_recipe(recipeCa) %>%
  add_model(gamma_spec) %>%
  fit(data = hapin_merged)

@@

#Fit the model
fitFe <- workflow() %>%
  add_recipe(recipeFe) %>%
  add_model(gamma_spec) %>%
  fit(data = hapin_merged)

@@

#Fit the model
fitK <- workflow() %>%
  add_recipe(recipeK) %>%
  add_model(gamma_spec) %>%
  fit(data = hapin_merged)

@@

#Fit the model
fitMg <- workflow() %>%
  add_recipe(recipeMg) %>%
  add_model(gamma_spec) %>%
  fit(data = hapin_merged)

@@

#Fit the model
fitMn <- workflow() %>%
  add_recipe(recipeMn) %>%
  add_model(gamma_spec) %>%
  fit(data = hapin_merged)

@@

#Fit the model
fitSul <- workflow() %>%
  add_recipe(recipeSul) %>%
  add_model(gamma_spec) %>%
  fit(data = hapin_merged)

@@

@@

#Fit the model
fitSi <- workflow() %>%
  add_recipe(recipeSi) %>%
  add_model(gamma_spec) %>%
  fit(data = hapin_merged)

@@

#Fit the model
fitTi <- workflow() %>%
  add_recipe(recipeTi) %>%
  add_model(gamma_spec) %>%
  fit(data = hapin_merged)

@@

#Fit the model
fitZn <- workflow() %>%
  add_recipe(recipeZn) %>%
  add_model(gamma_spec) %>%
  fit(data = hapin_merged)

@@

Based on the previous models, it seems like the most important variables are `fueltype` (as expected), `kerosene`, `stove_other` and `smoke` depending on the chemical species. I was worried of the low variability on responses for some of the variables so I am going to use only the variables `micro_exposures`, `macro_exposures`, `kerosene`, `stove_other` and `fueltype`.

```{r}
###########To predict the Al variable##############
# Create the recipe for preprocessing
recipeAl2 <- recipe(Al ~ micro_exposures + macro_exposures + stove_other + fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

#Fit the model
fitAl2 <- workflow() %>%
  add_recipe(recipeAl2) %>%
  add_model(gamma_spec) %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsAl2 <- fitAl2 %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsAl2)

###########To predict the BC variable##############
# Create the recipe for preprocessing
recipeBC2 <- recipe(BC ~ micro_exposures + macro_exposures + stove_other + fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

#Fit the model
fitBC2 <- workflow() %>%
  add_recipe(recipeBC2) %>%
  add_model(gamma_spec) %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsBC2 <- fitBC2 %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsBC2)

###########To predict the Ca variable##############
# Create the recipe for preprocessing
recipeCa2 <- recipe(Ca ~ micro_exposures + macro_exposures + stove_other + fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

#Fit the model
fitCa2 <- workflow() %>%
  add_recipe(recipeCa2) %>%
  add_model(gamma_spec) %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsCa2 <- fitCa2 %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsCa2)

###########To predict the Fe variable##############
# Create the recipe for preprocessing
recipeFe2 <- recipe(Fe ~ micro_exposures + macro_exposures + stove_other + fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

#Fit the model
fitFe2 <- workflow() %>%
  add_recipe(recipeFe2) %>%
  add_model(gamma_spec) %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsFe2 <- fitFe2 %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsFe2)

###########To predict the K variable##############
# Create the recipe for preprocessing
recipeK2 <- recipe(K ~ micro_exposures + macro_exposures + stove_other + fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

#Fit the model
fitK2 <- workflow() %>%
  add_recipe(recipeK2) %>%
  add_model(gamma_spec) %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsK2 <- fitK2 %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsK2)

###########To predict the Mg variable##############
# Create the recipe for preprocessing
recipeMg2 <- recipe(Mg ~ micro_exposures + macro_exposures + stove_other + fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

#Fit the model
fitMg2 <- workflow() %>%
  add_recipe(recipeMg2) %>%
  add_model(gamma_spec) %>%
  fit(data = hapin_merged)
